# Results

```{r}
library(openintro)
library(ggplot2)
library(tidyverse)
library(stringr)
library(agridat)
library(boot)
library(gridExtra)
library(ggridges)
library(broom)
library(reshape2)
library(GGally)
library(parcoords)
library(d3r)
library(Lock5withR)
library(grid)
library(vcd)
library(ggalluvial)
library(tidytext)
library(fmsb)
```

First, we will present an overview of the speed dating dataset by displaying the distribution of race, age, field of studies, and income of the participants across different genders. 

```{r}
sd <- df_select %>% 
  #filter(!is.na(race)) %>% 
  select(id_within_group,gender, race, age, income, field_study,race_importance,samerace,match)
sd <- sd[!duplicated(sd[ , 'id_within_group']), ] 
```

```{r}
sd %>% 
	group_by(race,gender) %>% 
	summarise(count = n()) %>% 
	ggplot(aes(x = reorder(race,(-count)), y = count)) + 
		geom_bar(stat = 'identity')+
    facet_wrap(~gender)+
  scale_x_discrete( labels = c('European', "Asian", "Hispanic",'Other','African American','Native American'))+
  coord_flip()+
  xlab('Race')+
  ylab('Count')
```
From the histogram above, we can observe that majority of the participants are White, and we barely see any Native American and African American participants. Hence, this dataset is very unbalanced in terms of diverse demographic background of participants. In addition to the distribution of race among participants, we also investigate how important is race from participants' perspectives. We generate the following histogram and density plot to demonstrate the importance of race.

```{r}
sd %>% 
  filter(!is.na(race_importance) & !race_importance==0) %>%
	group_by(race_importance,gender) %>% 
	summarise(count = n()) %>% 
	ggplot(aes(x = as_factor(race_importance),y=count)) + 
		geom_bar(stat = 'identity')+
    facet_wrap(~gender)+
  coord_flip()+
  xlab('Rating')+
  ylab('Count')
```

```{r}
sd %>% 
  filter(!is.na(race_importance) & !race_importance==0) %>%
	group_by(race_importance,gender) %>% 
	ggplot(aes(x = race_importance)) + 
		geom_density()+
    facet_wrap(~gender)+
  xlab('Rating')+
  ylab('Density')+
  scale_x_continuous(limits=c(0, 10),breaks = seq(0, 10, 1))
```
From the histograms and density diagrams above, we can observe that majority of male and female participants gave a low rating to the importance of race. Specifically, a larger proportion of men rated the importance of race as 1 than the female participants. Does lower rating in the importance of rating related to a higher rate of successful matching? Here, we calculated the success rate by dividing the number of positive responses (when matching=1) by the total number of matching results. Then, we generate a scatter plot to see if the importance of rate is related to the success rate of matching.
```{r}
df_copy<- df_select
pr <- df_copy %>% group_by(id_within_group) %>%   summarise(pos = sum(match))
tr <- df_copy %>% group_by(id_within_group) %>%    summarise(res=sum(n()))
temp= merge(as_tibble(pr),as_tibble(tr))
temp$SR=temp$pos/temp$res
temp<- merge(temp,df_select)
```

```{r}
temp<-temp[!duplicated(temp[ , 'id_within_group']),]

ggplot(temp,aes(x=race_importance,y=SR))+
  geom_point(alpha=0.6,color='blue')+
  scale_x_continuous(limits = c(0, 10), breaks = seq(0, 10, 1)) +
  scale_y_continuous(limits = c(0,1), breaks = seq(0, 1, 0.1))
```
We didn't see a salient relationship between success rate and the importance of race in the scatterplot above. However, we see that participants with high success rate (above 0.7) are those who rate importance of race with a score of 3 or below. Participants who rated importance of race with 10 face a much smaller group of ideal dating partners and consequently have a much lower rate of success.

We also generated a radar chart to compare the female and male ratings of the six attributes. We can observe similar features as in the boxplots. Male participants focus more on physical attractiveness than their female counterparts, while female participants value ambition more than the male participants. Male and female participants have similar perspectives on intelligence, sincerity and shared interests.
```{r}
df_t<-df_select
df_t <- df_t %>% 
  na.omit%>%
  mutate(sum1=  pre_rate_attractiveness+ pre_rate_ambition + pre_rate_sincerity + pre_rate_fun + pre_rate_share+pre_rate_intelligence) %>%
  mutate(pre_rate_attractiveness = (pre_rate_attractiveness/sum1)*100) %>% 
  mutate(pre_rate_sincerity = (pre_rate_sincerity/sum1)*100) %>% 
  mutate(pre_rate_intelligence= (pre_rate_intelligence/sum1)*100) %>% 
  mutate(pre_rate_ambition= (pre_rate_ambition/sum1)*100) %>% 
  mutate(pre_rate_share= (pre_rate_share/sum1)*100)%>%
  mutate(pre_rate_fun= (pre_rate_fun/sum1)*100)
feature<- df_t%>%
  select(id_within_group,gender,pre_rate_attractiveness:pre_rate_share )
male<-filter(feature,gender=='Male')
female<-filter(feature,gender=='Female')
column_label <- c("Attractive", "Sincere", "Intelligent", 
                  "Fun", "Ambitious", "Shared Interests")
row_label <- c("Male", "Female")
```

```{r}
radar_men <- as.data.frame(matrix(0, nrow = 2, ncol = 6))
colnames(radar_men) <- column_label
rownames(radar_men) <- row_label

for (i in (1:nrow(radar_men))) {
  for(j in c(1:ncol(radar_men))) {
    if(i == 1) {
      radar_men[i, j] <- mean(male[ , 2 + j])
    }
     if(i == 2){
      radar_men[i,j] <- mean(female[ , 2 + j])
    }  
  }
}
radar_men = rbind(rep(40, 5) , rep(0, 5) , radar_men)
radarchart(radar_men, pcol= c( rgb(0.2, 0.5, 0.5, 0.9), rgb(0.7, 0.5, 0.1, 0.9)), 
           pfcol = c(rgb(0.2, 0.5, 0.5, 0.4), rgb(0.7, 0.5, 0.1, 0.4)),  
           plwd = 3 , plty = 1, vlcex = 0.8, 
           title = "Attributes males find most important in Opposite Sex")
legend(x = 1, y = 1.2, legend = c("Male perspective", "Female perspective"), 
       bty = "n", pch = 20 , col = c(rgb(0.2, 0.5, 0.5, 0.4), rgb(0.7, 0.5, 0.1, 0.4)), 
       text.col = "black", cex = 0.8, pt.cex = 2)
```


```{r, fig.height = 5, fig.width=5}
df_traits <- df_select %>% na.omit() %>% select(id_within_group, gender, pre_rate_attractiveness, pre_rate_sincerity, pre_rate_intelligence, pre_rate_fun, pre_rate_ambition, pre_rate_share) %>% unique(by = id_within_group)
df_long <- melt(df_traits, id = c('id_within_group', 'gender')) %>% unite('merge', c(gender,variable), remove = TRUE)

#g1<- df_long %>% filter(genders == 'Female') %>% ggplot(aes(x = reorder(variable, value, median), y = value)) + #geom_boxplot() + scale_y_continuous(limits = c(0, 100))
#g2<- df_long %>% filter(genders == 'Male') %>% ggplot(aes(x = reorder(variable, value, median), y = value)) + geom_boxplot()  + scale_y_continuous(limits = c(0, 100))
#temp2<-list()
#temp2[[1]] = g1
#temp2[[2]] = g2
#grid.arrange(grobs = temp2, nrow = 2, ncol = 1)

df_long %>% ggplot(aes(x = reorder(merge, value, median), y = value)) + geom_boxplot()  + scale_y_continuous(limits = c(0, 100)) + coord_flip()

temp <- list()
t1 <- ggplot(df_traits, aes(x = pre_rate_attractiveness)) + geom_histogram(aes(y = ..density..),binwidth = 5) + xlim(0,100) + geom_density(color = 'red') + facet_wrap(~gender)
t4 <- ggplot(df_traits, aes(x = pre_rate_sincerity)) + geom_histogram(aes(y = ..density..),binwidth = 5) + xlim(0,100) + geom_density(color = 'red') + facet_wrap(~gender)
t2 <- ggplot(df_traits, aes(x = pre_rate_intelligence)) + geom_histogram(aes(y = ..density..),binwidth = 5) + xlim(0,100) + geom_density(color = 'red') + facet_wrap(~gender)
t3 <- ggplot(df_traits, aes(x = pre_rate_fun)) + geom_histogram(aes(y = ..density..),binwidth = 5) + xlim(0,100) + geom_density(color = 'red') + facet_wrap(~gender) 
t6 <- ggplot(df_traits, aes(x = pre_rate_ambition)) + geom_histogram(aes(y = ..density..),binwidth = 5) + xlim(0,100) + geom_density(color = 'red') + facet_wrap(~gender)
t5 <- ggplot(df_traits, aes(x = pre_rate_share)) + geom_histogram(aes(y = ..density..),binwidth = 5) + xlim(0,100) + geom_density(color = 'red') + facet_wrap(~gender)
temp[[1]] = t1
temp[[2]] = t2
temp[[3]] = t3
temp[[4]] = t4
temp[[5]] = t5
temp[[6]] = t6
grid.arrange(grobs = temp, nrow = 6, ncol = 1)
```

```{r, fig.height= 10, fig.width= 10}
df_temp <- df_select %>% na.omit() %>% select(samerace, match, race_importance) %>% mutate(IsSameRace = case_when(samerace == 1 ~ 'True', samerace == 0 ~ 'False'), IsMatch = case_when(match == 1 ~ 'True', match == 0 ~ 'False')) %>% select(-samerace, -match)

df_traits2 <- df_select %>% na.omit() %>% select(id_within_group, gender, decision, attractiveness, sincerity, intelligence, fun, ambition, shared_interest) %>% unique(by = id_within_group) %>% mutate(decision1 = case_when(decision == 1 ~ 'True', decision == 0 ~ 'False')) %>% select(-decision) %>% rename(decision = decision1) %>% filter(attractiveness == floor(attractiveness),sincerity == floor(sincerity),intelligence == floor(intelligence),fun == floor(fun),ambition == floor(ambition),shared_interest == floor(shared_interest))

mosaic(IsMatch ~ IsSameRace, direction = c("v", "h"), df_temp)
par(mfrow=c(2,3))
p1<-grid.grabExpr(mosaic(decision ~ attractiveness, direction = c("v", "h"), df_traits2))
p2<-grid.grabExpr(mosaic(decision ~ sincerity, direction = c("v", "h"), df_traits2))
p3<-grid.grabExpr(mosaic(decision ~ intelligence, direction = c("v", "h"), df_traits2))
p4<-grid.grabExpr(mosaic(decision ~ fun, direction = c("v", "h"), df_traits2))
p5<-grid.grabExpr(mosaic(decision ~ ambition, direction = c("v", "h"), df_traits2))
p6<-grid.grabExpr(mosaic(decision ~ shared_interest, direction = c("v", "h"), df_traits2))
grid.arrange(p1,p2,p3,p4,p5,p6,ncol=3,top="Main Title")
```