--- 
title: "SpeedDating Analysis"
author: "Sandy chen, Yixuan Liu"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
---
```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```

# Introduction

Do people with similar traits more likely to fall in love with each other? Do men and women have similar preferences in a romantic relationship? What are some most important attributes that contribute to a successful matching? Those are the questions at the heart of our investigation of the speed dating dataset. 

Speed dating nowadays becomes a common medium for establish romantic relationships. Speed dating allows participant to meet a group of potential romantic partners in a short amount of time. Disregard of the actual effectiveness of building a meaningful relationship within several minutes of dating, studying speed dating can answer some essential question about gender differences and racial differences in the dating market, which might shed some light on more complicated social issues. 


<!--chapter:end:index.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Data sources
The dataset was created and collected by two professors in the Columbia Business School, Raymond Fisman and Sheena S. Iyengar. The dataset was used for their paper [GENDER DIFFERENCES IN MATE SELECTION: EVIDENCE FROM A SPEED DATING EXPERIMENT](https://www8.gsb.columbia.edu/researcharchive/getpub/867/p). In this section, I will go over through the basic setting of this experiment and how they collected these data.

The data was collected during 2002 - 2004 speed dating experiments which is designed by those two professor. The data consists of 21 different experiment each of them has different number of participants, which is corresponded to the Wave feature in the data. In every experiment, every one would have a four minute conversation in an enclosed room setting with all other participants had different sex. After the first dating, they decided whether or not to meet someone again. These subjects are students from Columbia University. These participants were selected by answering some emails and fliers. In addition, they need complete a pre event survey which consists of age, field of study, race, tuition, zip code, income, goal, interests etc. In addition to these field, they were asked to rate six traits with respect to their potential dating partner (j=1) - Attractive, Sincere, Intelligent, Fun, Ambitious and Shared Interests. In addition, they were asked to rate what their opposite sex would look for in a date (j=2). Then, they were aksed to rate themselves (j=3) and how other perceived them (j=5) based on these six traits. Also, they were asked to rate how their same sex fellow think (j=4).

After each round (4 minute dating), they had a scorecard to fill to show their likeness and willingness of a future dating (Yes or No) and rate their dating subject based on the six traits. In the half way of each experiment, these participants are asked to fill another questionarie to rate the importance of six traits in a potential date and their opinion of themselves.

After one day of meeting, these pariticpants were aksed to fill another survey which consists same questions as the pre survey and to rate six traits.

Finally, After three or four week, they were asked again to rate six traits.

So as a result, the dataset consist of fews survey which has format attr j_i. The corresponding j values are stated above. In addition, i = 1 means pre-survey time and i = 2 means day 1 survey time and i = 3 means three or four week survey time. During the event, their rating has no j_i follow attr: For example, sinc means how this person rate their partner in sincere after 4 mins meeting. In addition, attr_o means the rate that their partner give after meeting.

One problem with these dataset is that for each person, there are multiple rows and most of their column are same. The difference correspond to different partner's rating and their own rating to different partners. The more specific description of each variable can be found [here](https://github.com/yixual5/speedDating/blob/main/Speed%20Dating%20Data%20Key.doc).


<!--chapter:end:02-data.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Data transformation

<!--chapter:end:03-cleaning.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Missing values
```{r}
## Read in the dataset
dating_df=read.csv('Speed Dating Data.csv')
```

## Missing Value Pattern Analysis: Whole Dataset

In this section, we will visualize and analyze the missing values in the dataset. The dataset we use contains two part: one part are variables documented by the event, and the rest of the data came from survey results filled out voluntarily by the participants of the speed dating events. In order to analyze the missing value trend, we are first going to generate a histogram of the percentage of missing value of all variables, which hopefully will give us some insight for the investigation of the dataset. Due to a large amount of variables, we cannot plot all of them and we decide to see some of important variables in the pattern plot. In later section, we try to remove some unnecessary varaibles and some variables missing a lot of values.

Below, we calculate the missing value percentage by dividing the number of the null values by the total number of entries in each column. 
```{r}
missing<- sapply(dating_df, function(x) sum(is.na(x))/length(x))
missing1<-as.data.frame(missing)
missing1$var=rownames(missing1)
rownames(missing1)<-NULL
```

We would plot the missing datta percentage with a histogram below:

```{r,fig.height=10,fig.width=10}
## Histogram:
ggplot(missing1,aes(fct_reorder(as.character(var),missing),y=missing))+
  geom_col()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  ylab('Percent of Missing Value')+
  xlab('Variable')

```

From the histogram above, we can observe that variables like id, iid, wave number, and matching results which are data documented by the event have no missing values. Variables like race, career, and zipcode, and gender also contain no missing values. However, variables involving the evaluation of partners during the event, for example, fun and sinc (sincerity) have missing value exceeding 50%. It will be interesting to look into the reason behind the high missing percentage of those variables. For instance, different waves of the participants were asked to fill out different evaluative questions, and hence variables related to evaluation contain a large portion of missing values. In addition, it is kind of hard to answer these questionnaires with a lot of subjective questions.


## Missing Value Pattern Analysis: Subset of Variables 

Based on the questions we want to investigate, we select 15 variables among the 195 variables from the dataset. Below, we plot the missing value pattern for these 15 variables: 

```{r}
df_select<-dating_df[,c('gender','match','race_o','field','race','income','goal','age','attr','dec_o','like_o','imprace','mn_sat','attr1_1','met')]

```


```{r}
### Draw missing plots:
missing_plot <- function(x, percent) {
  missing_patterns <- data.frame(is.na(x)) %>%
  group_by_all() %>%
  count(name = "count", sort = TRUE) %>%
  ungroup()
  lv <- data.frame(is.na(x)) %>% gather() %>% group_by(key) %>%  summarise(value = sum(value==TRUE)) %>% arrange(.,desc(value))
  temp <-t(apply(missing_patterns[,names(missing_patterns) != "count"] , 1, FUN = function(x, output) {
      t = FALSE
   for (i in 1 : length(x)) {
     t = t || x[i]
     if (x[i]) {
       x[i] = 1
     } else {
       x[i] = 0
     }
   }
    if (!t) {
      for (i in 1 : length(x)) {
        x[i] = 2
      }
    }
    return (x)
  })) %>% melt()
  
  p1 <- ggplot(temp, aes(x = fct_relevel(Var2, list(as.character(lv$key))), y = fct_rev(as.factor(Var1)) , fill = as.character(value))) + geom_tile(color = 'white') +
       xlab("Variables") + ylab("Missing Pattern") + theme(legend.position = "None") + scale_fill_manual(values = c("grey", "#CF9FFF", "#A9A9A9")) + scale_x_discrete(label = abbreviate)
  if (percent) {
    p2 <- ggplot(lv, aes(x = fct_relevel(key, list(as.character(key))), y = 100 * value / sum(missing_patterns$count))) + geom_col() + xlab("") + ylab("% rows:\n missing:") + ylim(0,100) + scale_x_discrete(label = abbreviate)
  } else {
    p2 <- ggplot(lv, aes(x = fct_relevel(key, list(as.character(key))), y = value)) + geom_col() + xlab("") + ylab("num rows:\n missing:") + scale_x_discrete(label = abbreviate)
  }
  if (percent) {
    p3 <- rownames_to_column(missing_patterns, var = "ids") %>% ggplot(aes(fct_relevel(ids, rev(c(ids))), 100 * count / sum(count))) + geom_col() + xlab("") + ylab("% row") + coord_flip() + ylim(0,100)
  } else {
    temp2 <- rownames_to_column(missing_patterns, var = "ids")
    p3 <- temp2 %>% ggplot(aes(fct_relevel(ids, rev(c(ids))), count)) + geom_col() + xlab("") + ylab("row count") + coord_flip()
  }
  
  p4 <- (p2 / p1) + plot_layout(height = c(1,4))
  p5 <- (plot_spacer() / p3) + plot_layout(height = c(1,4), width = 0.1)
  p2 + plot_spacer() +
    p1 + p3 +
    plot_layout(width = c(4,1), height = c(1,4))
}
```

```{r}
missing_plot(df_select,FALSE)
missing_plot(df_select,TRUE)
```

From the plots above, we can observe that Variables such as field, gender, income, match, and decision after the meeting have no missing values. We can observe that in the dataset, variables age, a1_1(importance of attractiveness), goal, impr (importance of same race) are often missing together. We need to further investigate the high missing value percentage of met (if youm met your partner before), and the result might give us some interesting fact about the dataset. 

<!--chapter:end:04-missing.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Results

<!--chapter:end:05-results.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Interactive component



<!--chapter:end:06-interactive.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Conclusion


<!--chapter:end:07-conclusion.Rmd-->

